---
title: "REDCap Automated Billing Revenue, Status, and Projections"
author: "Philip Chase"
format:
  html:
    code-fold: true
    code-tools: true
    embed-resources: true
    df-print: kable
    fig-width: 9
    fig-height: 3.5
  # typst:
  #   papersize: a3
  #   margin:
  #     x: 1.9cm
  #     y: 1.9cm
  #   columns: 1
  #   df-print: kable
  #   fig-width: 9
  #   fig-height: 3.5
  # pdf:
  #   geometry:
  #     - margin=19mm
  #     - bottom=30mm
  #   fig-pos: 'H'
  #   df-print: kable
editor: visual
date: "`r Sys.Date()`"
---

The CTS-IT manages the REDCap system for the UF Clinical Translational Science Institute (CTSI). CTS-IT is a business within the University of Florida called a *UF Auxillary*. As such, CTS-IT charges for its services to cover the cost of providing those services. CTS-IT charges its customers annually for the REDCap projects they use and for time spent managing individual REDCap projects on behalf of the study team. CTS-IT also does contractual work--often involving REDCap--for studies based on a written agreement. These charges are assessed monthly be a system called *REDCap Automated Billing*. This report describes the revenue generated by this system assessing the current revenue, estimating future revenue, and providing other metrics of growth and performance.

While the original billing system started in July 2018, that system was rewritten from the ground up in 2022. The revision focuses on automated processes to implement a monthly billing cycle and delegates the billing work to the UF CTSI Service Billing Team. The new system was implemented in September 2022. In Fall 2024, CTS-IT added billing for REDCap support for work and contractual work.

```{r}
#| label: calcs
#| output: false
#| echo: false
#| warning: false
library(redcapcustodian)
library(rcc.billing)
library(RMariaDB)
library(DBI)
library(tidyverse)
library(lubridate)
library(dotenv)
library(fs)
library(scales)
library(gt)
library(ggplot2)

if (Sys.getenv("RCCBILLING_DB_NAME") == "") dotenv::load_dot_env(here::here("prod.env"))

init_etl("revenue_status_and_projections")

rcc_billing_conn <- connect_to_rcc_billing_db()

invoice_status <-
  tbl(rcc_billing_conn, "invoice_line_item") %>%
  collect() %>%
  mutate(invoice_month = lubridate::floor_date(created - dmonths(1), unit = "month")) %>%
  filter(status %in% c("invoiced", "paid"))

revenue <-
  tbl(rcc_billing_conn, "invoice_line_item") %>%
  collect() %>%
  mutate(invoice_month = lubridate::floor_date(created - dmonths(1), unit = "month")) %>%
  filter(status != "canceled") %>%
  group_by(status, invoice_month) %>%
  summarise(total = sum(amount_due)) %>%
  tidyr::pivot_wider(
    id_cols = c("invoice_month"),
    names_from = status,
    values_from = total
  ) %>%
  mutate(
    invoices_sent = coalesce(invoiced, 0) + coalesce(paid, 0),
    portion_paid = coalesce(paid, 0) / invoices_sent
  ) |>
  ungroup() |>
  arrange(invoice_month)

# The average portion paid should ignore the last three months as sequester_unpaid_projects has not yet been applied to these months
average_portion_paid <- revenue %>%
  slice_head(n = nrow(revenue) - 4) %>%
  summarise(average_portion_paid = mean(portion_paid)) %>%
  pull(average_portion_paid)

expected_revenue <- revenue %>%
  mutate(expected_revenue = case_when(
    row_number() <= nrow(revenue) - 4 ~ paid,
    TRUE ~ round(coalesce(invoices_sent, sent) * average_portion_paid)
  ))

revenue_and_projections <- bind_cols(
  expected_revenue %>%
    summarise(
      actual_revenue = sum(paid, na.rm = T),
      accounts_receivable = sum(invoiced, na.rm = T)
    ),
  # Base expected revenue only on the last 12 months of revenue
  expected_revenue |>
    slice_tail(n = 12) |>
    # correct expected revenue for the rate increase that started 2023-11-01
    mutate(expected_revenue = if_else(invoice_month < ymd("2023-10-15"), expected_revenue * 1.30, expected_revenue)) |>
    summarise(
      expected_monthly_revenue = round(mean(expected_revenue, na.rm = T)),
      expected_annual_revenue = 12 * expected_monthly_revenue
    )
)
```

```{r}
#| label: service-type-data-prep
#| echo: false
#| warning: false

## Identify contractual work
source_credentials <- get_redcap_credentials(Sys.getenv("REDCAP_SERVICE_REQUEST_PID"))

# Get raw Service Request data
fields_to_fetch <- c("record_id", "study_name", "study_complete")
service_requests <- REDCapR::redcap_read(
  redcap_uri = source_credentials$redcap_uri,
  token = source_credentials$token,
  batch_size = 2000,
  fields = fields_to_fetch
)$data

contract_support <- service_requests |>
  filter(!is.na(study_name)) |>
  select(any_of(fields_to_fetch)) |>
  mutate(contractual = T,
         record_id = as.character(record_id),
         service_type_code = 2
  )

service_type_data <- tbl(rcc_billing_conn, "service_type") |>
  collect() |>
  bind_rows(
    tribble(
      ~service_type_code, ~service_type, ~price, ~billing_frequency,
      99, "Contractual", 130, 0
    )
  ) |>
  mutate(service_type = if_else(service_type_code == 2, "Support", service_type))

invoice_line_items_with_fy_and_month_paid <- invoice_status |>
  # describe invoice_line_items by the month in which they were paid
  filter(status == "paid") |>
  mutate(
    Status = "Paid",
    `Month Received` = as_date(floor_date(je_posting_date, unit = "month"))
  ) |>
  # Get the Fiscal year in which the invoice was paid
    mutate(month_received = lubridate::as_datetime(
    `Month Received`,
    tz = Sys.getenv("TIME_ZONE")
  ) + days(1)) |>
  cross_join(rcc.billing::fiscal_years) |>
  filter(month_received %within% fy_interval) |>
  rename(FY = csbt_label) |>
  select(-c(Status, fy_interval, month_received))
```

```{r}
#| label: revenue-by-service-type
#| echo: false
#| warning: false

revenue_by_service_type <-
  invoice_line_items_with_fy_and_month_paid |>
  # Mutate service_type_code here to get 99s on contractual work
  left_join(contract_support |> select(record_id, contractual, service_type_code),
    by = c("service_identifier" = "record_id", "service_type_code")
  ) |>
  mutate(service_type_code = case_when(
    contractual ~ 99,
    T ~ service_type_code
  )) |>
  mutate(service_type = factor(
    x = service_type_code,
    levels = service_type_data$service_type_code,
    labels = service_type_data$service_type
  )) |>
  group_by(FY, service_type) |>
  summarise(revenue = sum(amount_due, na.rm = T)) |>
  ungroup() |>
  pivot_wider(
    id_cols = "FY",
    names_from = "service_type",
    values_from = "revenue"
  ) |>
  select(
    `Fiscal Year` = FY,
    "Annual REDCap Project Maintenance",
    "Support",
    "Contractual"
  )

current_fy_months <- 
  invoice_line_items_with_fy_and_month_paid |>
    group_by(FY, `Month Received`) |>
    summarize(qty_invoices = n()) |>
    add_count(FY) |>
    ungroup() |>
    filter(`Month Received` == max(`Month Received`)) |>
  pull(n)
  
```

```{r}
#| label: invoicing-by-service-type
#| echo: false
#| warning: false

invoicing_by_service_type <-
  tbl(rcc_billing_conn, "invoice_line_item") |>
  collect() |>
  # Mutate service_type_code here to get 99s on contractual work
  left_join(contract_support |> select(record_id, contractual, service_type_code),
    by = c("service_identifier" = "record_id", "service_type_code")
  ) |>
  mutate(service_type_code = case_when(
    contractual ~ 99,
    T ~ service_type_code)) |>
  mutate(service_type = factor(x = service_type_code,
       levels = service_type_data$service_type_code,
       labels = service_type_data$service_type)) |>
  mutate(invoice_month = as.Date(lubridate::floor_date(created - dmonths(1), unit = "month"))) |>
  filter(!status %in% c("canceled", "sent")) |>
   group_by(status, invoice_month, service_type) |>
  summarise(total = sum(amount_due, na.rm = T)) |>
  tidyr::pivot_wider(
    id_cols = c("invoice_month", "service_type"),
    names_from = status,
    values_from = total
  ) |>
  mutate(
    invoices_sent = coalesce(invoiced, 0) + coalesce(paid, 0)
  ) |>
  ungroup() |>
  arrange(invoice_month)
```

## Revenue and Receivables

To date, the new system has generated `r scales::dollar(revenue_and_projections$actual_revenue)`. The current Accounts Receivable is `r scales::dollar(revenue_and_projections$accounts_receivable)`. Revenue is broken into three categories: Annual REDCap Project Maintenance, Support, and Contractual. *Annual REDCap Project Maintenance* is the fee charged for keep a REDCap project online for one year. *Support* is adhoc work done to maintain specific projects. *Contractual* is work based on a written agreement. Revenue for these categories by fiscal year is shown in @fig-revenue-by-service-type

```{r}
#| label: fig-revenue-by-service-type
#| fig-cap: Annual REDCap billing revenue by service
#| echo: false
#| warning: false

revenue_by_service_type |>
  gt() |>
  tab_header(
    title = "Revenue by Fiscal Year and Service",
  ) |>
  fmt_currency(decimals = 0) |>
  tab_footnote(
    footnote = glue::glue("Approximately {current_fy_months} of 12 months of data are included in these metrics."),
    locations = cells_body(
      columns = "Fiscal Year",
      rows = `Fiscal Year` == max(`Fiscal Year`)
    ),
    placement = "right"
  )
```

Monthly invoicing by Service is shown in @fig-invoicing-by-service-type

```{r}
#| label: fig-invoicing-by-service-type
#| fig-cap: "Invoicing by service since REDCap Automated billing began"
#| fig-height: 5
#| echo: false
#| warning: false

invoicing_by_service_type |>
  rename(`Month Received` = invoice_month,
         Service = service_type) |>
  ggplot(mapping = aes(x = `Month Received`, fill = Service, y = invoices_sent)) +
  facet_wrap(facets = "Service", ncol = 1) +
  geom_bar(position = "stack", stat = "identity") +
  scale_y_continuous(
    labels = dollar
  ) +
  # geom_smooth(method = "lm") +
  scale_x_date(date_breaks = "3 month", date_labels = "%Y-%m") +
  theme(axis.text.x = element_text(angle = 30)) +
  theme(legend.position="bottom") +
  xlab("Month Received") +
  ylab("Amount invoiced")
```

## Pro Bono work

CTS-IT does not charge for all its services. It offers a free consultation of up to an hour for every REDCap project on the CTSI REDCap system. This allows our customers to get free REDCap expertise as they plan their data collection efforts. Similarly, CTS-IT forgives up to a year of charges for every REDCap project if the owner deletes it before the next invoicing. This allows study teams to learn REDCap for free, taking tutorials, building projects, testing designs, and building the data collection instruments they will ultimately use in their study. Studies can be built, deployed, and completed entirely free of charge if done in under one year.

Table @probono-summary-table shows the value of these services by fiscal year since CTS-IT started carefully accounting for charged and non-charged services. Figure @fig-probono-faceted shows the monthly detail underlying the annual figures. Support billing started 21 months after the annual project billing.

```{r}
#| label: probono-support-prep
#| echo: false
#| warning: false

# Support billing - Pro Bono
probono_support <- tbl(rcc_billing_conn, "invoice_line_item") |>
  filter(service_type_code == 2, price_of_service == 0) |>
  select(id, qty_provided, fiscal_year, month_invoiced, created) |>
  collect()

normal_service_rate <- tbl(rcc_billing_conn, "invoice_line_item") |>
  filter(service_type_code == 2, price_of_service > 0) |>
  select(fiscal_year, month_invoiced, price_of_service) |>
  group_by(fiscal_year, month_invoiced, price_of_service) |>
  count(fiscal_year, month_invoiced, price_of_service) |>
  ungroup() |>
  collect() |>
  mutate(old_order = row_number()) |>
  arrange(desc(n)) |>
  distinct(fiscal_year, month_invoiced, .keep_all = T) |>
  arrange(old_order) |>
  select(-c(old_order, n))

probono_support_with_rate <- probono_support |>
  left_join(normal_service_rate, by = c("fiscal_year", "month_invoiced")) |>
  mutate(value = qty_provided * price_of_service) |>
  select(fiscal_year, month_invoiced, value) |>
  mutate(revenue_source = "Support")
```

```{r}
#| label: apb-probono
#| echo: false
#| warning: false

rc_conn <- connect_to_redcap_db()

redcap_projects <- dplyr::tbl(rc_conn, "redcap_projects") |>
  collect()

project_details <- get_project_details_for_billing(
  rc_conn, rcc_billing_conn, redcap_projects$project_id)|>
  select(project_id, billable)

project_history <- tbl(rcc_billing_conn, "redcap_projects_history")|>
  select(project_id, project_creation_time = creation_time, 
         date_deleted)|>
  collect()

project_history <- project_history|>
  left_join(project_details, by = "project_id")

apb_service_rate <- tbl(rcc_billing_conn, "invoice_line_item") |>
  filter(service_type_code == 1, price_of_service > 0) |>
  select(fiscal_year, month_invoiced, price_of_service) |>
  group_by(fiscal_year, month_invoiced, price_of_service) |>
  count(fiscal_year, month_invoiced, price_of_service) |>
  ungroup() |>
  collect() |>
  mutate(old_order = row_number()) |>
  arrange(desc(n)) |>
  distinct(fiscal_year, month_invoiced, .keep_all = T) |>
  arrange(old_order) |>
  select(-c(old_order, n))

probono_apb <- project_history |>
  filter(billable == 1, !is.na(date_deleted), date_deleted >= as.Date("2022-09-07")) |>
  mutate(
    days_active = as.numeric(as.Date(date_deleted) - as.Date(project_creation_time)),
    month_invoiced = format(as.Date(date_deleted), "%m"),
    fiscal_year = map_chr(date_deleted, ~ rcc.billing::fiscal_years |> filter(as.Date(.x) %within% fy_interval) |> pull(csbt_label))
  ) |>
  left_join(apb_service_rate, by = c("fiscal_year", "month_invoiced")) |>
  mutate(
    value = (days_active / 365) * price_of_service
  ) |>
  select(fiscal_year, month_invoiced, value) |>
  mutate(revenue_source = "Annual Project Billing")
```

```{r}
#| label: combine-support-apb
#| echo: false
#| warning: false

# Combine support and APB probono
probono_combined <- bind_rows(probono_support_with_rate, probono_apb) |>
  filter(!is.na(fiscal_year), !is.na(month_invoiced), !is.na(value))

# Summarize by FY and by month
probono_by_fy <- probono_combined |>
  group_by(fiscal_year) |>
  summarise(`Value of Pro Bono Services` = sum(value, na.rm = TRUE)) |>
  ungroup()

probono_by_month <- probono_combined |>
  group_by(revenue_source, fiscal_year, month_invoiced) |>
  summarise(`Value` = sum(value, na.rm = TRUE)) |>
  ungroup()
```

```{r}
#| label: probono-summary-table
#| echo: false
#| warning: false

probono_by_fy <- probono_combined |>
  group_by(fiscal_year) |>
  summarise(`Value of Pro Bono Services` = sum(value, na.rm = TRUE), .groups = "drop")

invoice_line_items_with_fy_and_month_paid <- invoice_status |>
  filter(status == "paid") |>
  mutate(
    `Month Received` = as.Date(lubridate::floor_date(je_posting_date, unit = "month"))
  ) |>
  mutate(month_received = lubridate::as_datetime(
    `Month Received`,
    tz = Sys.getenv("TIME_ZONE")
  ) + lubridate::days(1)) |>
  dplyr::left_join(rcc.billing::fiscal_years, by = character()) |>
  filter(month_received %within% fy_interval) |>
  rename(FY = csbt_label) |>
  select(-c(fy_interval, month_received))

revenue_by_service_type <- invoice_line_items_with_fy_and_month_paid |>
  left_join(contract_support |> select(record_id, contractual, service_type_code),
            by = c("service_identifier" = "record_id", "service_type_code")) |>
  mutate(service_type_code = dplyr::case_when(
    contractual ~ 99,
    TRUE ~ service_type_code
  )) |>
  mutate(service_type = factor(
    x = service_type_code,
    levels = service_type_data$service_type_code,
    labels = service_type_data$service_type
  )) |>
  group_by(FY, service_type) |>
  summarise(revenue = sum(amount_due, na.rm = TRUE), .groups = "drop") |>
  tidyr::pivot_wider(
    id_cols = "FY",
    names_from = "service_type",
    values_from = "revenue"
  ) |>
  select(
    `Fiscal Year` = FY,
    "Annual REDCap Project Maintenance",
    "Support",
    "Contractual"
  )

revenue_by_fy <- revenue_by_service_type |>
  mutate(
    Revenue = rowSums(across(-`Fiscal Year`), na.rm = TRUE)
  ) |>
  select(FY = `Fiscal Year`, Revenue)

probono_summary <- revenue_by_fy |>
  left_join(probono_by_fy, by = c("FY" = "fiscal_year")) |>
  mutate(
    `Pro Bono Services as a portion of revenue` = `Value of Pro Bono Services` / Revenue
  ) |>
  select(
    FY,
    Revenue,
    `Value of Pro Bono Services`,
    `Pro Bono Services as a portion of revenue`
  )

gt::gt(probono_summary) |>
  gt::fmt_currency(columns = c("Revenue", "Value of Pro Bono Services"), decimals = 0) |>
  gt::fmt_percent(columns = "Pro Bono Services as a portion of revenue", decimals = 0) |>
  gt::tab_header(
    title = "Pro Bono Value and Portion of Revenue by Fiscal Year"
  )
```

```{r}
#| label: fig-probono-faceted
#| fig-cap: "Pro Bono value by month and source"
#| echo: false
#| warning: false
#| fig-height: 5

probono_combined |>
  mutate(
    month_num = match(month_invoiced, month.name),
    date = as.Date(paste0(str_sub(fiscal_year, 1, 4), "-", sprintf("%02d", month_num), "-01"))
  ) |>
  filter(!is.na(date)) |>
  group_by(revenue_source, date) |>
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = date, y = value, fill = revenue_source)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~revenue_source, scales = "free_y") +
  scale_y_continuous(labels = dollar) +
  scale_x_date(date_breaks = "3 month", date_labels = "%Y-%m") +
  labs(
    x = "Month",
    y = "Pro Bono Value",
    title = "Pro Bono Value by Month and Source"
  ) +
  theme(axis.text.x = element_text(angle = 30))
```

## Aging Report

The unpaid invoices breakdown as shown in @fig-aging-report. For invoices more than 4 months old, the payment rate is `r scales::percent(average_portion_paid)`.

```{r}
#| label: fig-aging-report
#| fig-cap: Unpaid invoices grouped by age
#| echo: false
#| warning: false

invoice_status %>%
  filter(status == "invoiced") %>%
  select(invoice_month, status, amount_due, created) %>%
  mutate(age_in_days = now() - created) %>%
  mutate(`Aging Group` = as_factor(case_when(
    age_in_days <= 30 ~ "0-30 days",
    age_in_days <= 60 ~ "31-60 days",
    age_in_days <= 90 ~ "61-90 days",
    age_in_days <= 120 ~ "91-120 days",
    age_in_days <= 150 ~ "121-150 days",
    age_in_days <= 180 ~ "151-180 days",
    TRUE ~ "180+ days"
  ))) %>%
  count(`Aging Group`, name = "Invoices") %>%
  arrange(desc(`Aging Group`)) |>
  gt()
```

## Contractual invoicing

@fig-recent-contractual shows the Contractual revenue generated in the past 3 months

```{r}
#| label: contract_intermediate_calcs
#| echo: false
#| warning: false
#| include: false

invoiced_payable_consulting_with_invoice_month <-
  tbl(rcc_billing_conn, "invoice_line_item") |>
    filter(service_identifier %in% contract_support$record_id) |>
    filter(service_type_code == 2) |>
    filter(price_of_service > 0) |>
    filter(!status %in% c("canceled", "sent")) |>
    collect() |>
    mutate(invoice_month = as.Date(lubridate::floor_date(created - dmonths(1), unit = "month")))

start_month <- as.Date(lubridate::floor_date(today() - dmonths(4), unit = "month"))
end_month <- invoiced_payable_consulting_with_invoice_month |> summarise(end_month = max(invoice_month)) |>
  pull(end_month)

```

```{r}
#| label: fig-recent-contractual
#| fig-cap: "Recent Contractual invoices by invoiced amount"
#| echo: false
#| warning: false

  invoiced_payable_consulting_with_invoice_month |>
    left_join(contract_support |> select(record_id, study_name), by = c("service_identifier" = "record_id")) |>
    # get the last 3 months of Contractual invoices
    filter(invoice_month >= start_month) |>
    group_by(status, invoice_month, study_name) |>
    summarise(total = sum(amount_due, na.rm = T)) |>
    pivot_wider(
      id_cols = c("invoice_month", "study_name"),
      names_from = "status",
      values_from = "total"
    ) |>
    ungroup() |>
    arrange(invoice_month, desc(coalesce(paid, invoiced))) |>
    gt(groupname_col = "invoice_month", rowname_col = "study_name") |>
    tab_header(
      title = "Recent Contractual invoicing and revenue",
      subtitle = glue::glue("Service months: {start_month} - {end_month}")
    ) |>
    fmt_currency() |>
    tab_stub_indent(
      rows = everything(),
      indent = 3
    )
```
