---
title: "cleanup_bad_email_addresses"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cleanup_bad_email_addresses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(rcc.billing)
library(DiagrammeR)
```

Note: this script can no longer read mailboxes. This started in October 2022 when UF moved its Office 365 service to OAuth2. The script will throw warnings about this issue, but it still runs.

Locate bad UF addresses in REDCap replace them if possible, erase them if not, then disable accounts with no primary email address. 

This script reads the mail of a dummy in box that is used as the sender of numerous automated emails. When those messages generate bounce messages, they are delivered to the dummy inbox. This script reads that inbox, searches for message subject lines it recognizes and parses the message body to locate bad email addresses. It searches UF person data for a suitable replacement email address. If it finds one it replaces the bad address where it occurs. Otherwise it erases the bad email address from REDCap.

```{r mermaid_diagram, echo=FALSE, fig.align='center'}
DiagrammeR::mermaid("graph TD
    subgraph Suppliers
        S1[REDCap Database]
        S2[Email System]
        S3[RCC Billing Database]
        S4[VIVO Database]
    end

    subgraph Inputs
        I1[REDCap User Information]
        I2[Bounced Email Messages]
        I3[RCC Job Log]
        I4[UF Person Data]
    end

    subgraph Process
        P1[Read Dummy Inbox]
        P2[Parse Message Body]
        P3[Locate Bad UF Addresses]
        P4[Search for Replacement]
        P5[Update or Erase Address]
        P6[Disable Accounts]
    end

    subgraph Outputs
        O1[Updated REDCap User Information]
        O2[Updated RCC Job Log]
    end

    subgraph Customers
        C1[REDCap Users]
        C2[RCC Billing System]
    end

    S1 --> I1
    S2 --> I2
    S3 --> I3
    S4 --> I4

    I1 --> P1
    I2 --> P1
    P1 --> P2
    P2 --> P3
    I3 --> P3
    I4 --> P4
    P3 --> P4
    P4 --> P5
    P5 --> P6

    P6 --> O1
    P6 --> O2

    O1 --> C1
    O2 --> C2
    ", width = 700, height= 700)

```
