---
title: "cleanup_project_ownership_table"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cleanup_project_ownership_table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(rcc.billing)
library(DiagrammeR)
```

`cleanup_project_ownership_table.R` identifies projects with ownership issues, nominates a new owner if possible, and write that owner to the project ownership data. It cannot always determine a new owner. It logs the changes it makes to ownership data and the projects where it cannot fix the ownership data.
```{r mermaid_diagram, echo=FALSE, fig.width=8, fig.align='center'}
DiagrammeR::mermaid("
graph LR
    subgraph Suppliers
        S1[REDCap Database]
    end

    subgraph Inputs
        I1[redcap_entity_project_ownership]
        I2[redcap_user_information]
        I3[redcap_projects]
        I4[redcap_user_rights]
        I5[redcap_user_roles]
    end

    subgraph Process
        P1[Identify Projects with Ownership Issues]
        P2[Get Projects Needing New Owners]
        P3[Get Projects Without Owners]
        P4[Get Research Projects Not Using Viable PI Data]
        P5[Nominate New Owners]
        P6[Update Project Ownership Data]
        P7[Blank Unresolvable Ownership Records]
        P8[Log Changes and Issues]

        P1 --> P2
        P1 --> P3
        P1 --> P4
        P2 --> P5
        P3 --> P5
        P4 --> P5
        P5 --> P6
        P5 --> P7
        P6 --> P8
        P7 --> P8
    end

    subgraph Outputs
        O1[Updated redcap_entity_project_ownership]
        O2[rcc_job_log]
    end

    subgraph Customers
        C1[REDCap Database]
        C2[RCC Billing System]
    end

    S1 --> I1
    S1 --> I2
    S1 --> I3
    S1 --> I4
    S1 --> I5

    I1 --> P1
    I2 --> P1
    I3 --> P1
    I4 --> P1
    I5 --> P1

    P8 --> O1
    P8 --> O2

    O1 --> C1
    O2 --> C2
")
```
# Sources
- redcap database tables

# Inputs
 - redcap database tables:
   - redcap_entity_project_ownership
   - redcap_user_information
   - redcap_projects
   - redcap_user_rights
   - redcap_user_roles

# Processes
 - Identify projects with ownership issues, nominate a new owner if possible, and write that owner to the ownership data.

# Outputs
 - redcap_entity_project_ownership
    - blanked or revised project_ownership records
 - rcc_job_log
    - redcap_entity_project_ownership update_records
    - redcap_entity_project_ownership insert_records
    - project_ownership_records_with_issues_and_no_updates
 
# Consumers
- redcap database tables: redcap_entity_project_ownership
- rcc billing tables: rcc_job_log
