---
title: "cleanup_bad_project_pi_email_addresses"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cleanup_bad_project_pi_email_addresses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(rcc.billing)
library(DiagrammeR)
```

Note: this script can no longer read mailboxes. This started in October 2022 when UF moved its Office 365 service to OAuth2. The script will throw warnings about this issue, but it still runs.

Locate bad email addresses in Project PI data, replace them if possible, erase them if not.

This script reads the mail of a dummy in box that is used as the sender of numerous automated emails. When those messages generate bounce messages, they are delivered to the dummy inbox. This script reads that inbox, searches for message subject lines it recognizes and parses the message body to locate bad email addresses. It searches REDCap user information for a suitable replacement email address. If it finds one it replaces the bad address where it occurs. Otherwise it erases the bad email address from Project PI data.

```{r mermaid_diagram, echo=FALSE, fig.align='center'}
DiagrammeR::mermaid("
graph TB
    subgraph Suppliers
        S1[REDCap Database]
        S3[RCC Billing Database]
        S2[Email System]
    end

    subgraph Inputs
        I1[REDCap User Information]
        I4[REDCap Projects]
        I3[RCC Job Log]
        I2[Bounced Email Messages]
    end

    subgraph Process
        P1[Read dummy sender's inbox]
        P2[Parse message body for rejects]
        P3[Locate bad PI email <br/>addresses in project data]
        P4[Search REDCap user data for a replacement]
        P5[Update or erase address]
        P6[Read logs of email sending]
        P7[Parse logs for rejects]
    end

    subgraph Outputs
        O1[Updated REDCap Projects]
        O2[Updated RCC Job Log]
    end

    subgraph Customers
        C1[REDCap Users]
        C2[RCC Billing System]
    end

    S1 --> I1
    S1 --> I4
    S2 --> I2
    S3 --> I3

    I1 --> P4
    I2 --> P1
    P1 --> P2
    P2 --> P3
    I4 --> P3
    P3 --> P4
    P4 --> P5
    I3 --> P6
    P6 --> P7
    P7 --> P3

    P5 --> O1
    P5 --> O2

    O1 --> C1
    O2 --> C2
", width = 700, height= 700)
```
