---
title: "cleanup_bad_email_addresses"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cleanup_bad_email_addresses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rcc.billing)
```

Note: this script can no longer read mailboxes. This started in October 2022 when UF moved its Office 365 service to OAuth2. The script will throw warnings about this issue, but it still runs.

Locate bad UF addresses in REDCap replace them if possible, erase them if not, then disable accounts with no primary email address. 

This script reads the mail of a dummy in box that is used as the sender of numerous automated emails. When those messages generate bounce messages, they are delivered to the dummy inbox. This script reads that inbox, searches for message subject lines it recognizes and parses the message body to locate bad email addresses. It searches UF person data for a suitable replacement email address. If it finds one it replaces the bad address where it occurs. Otherwise it erases the bad email address from REDCap.

_TODO: Insert a SIPOC diagram here. Create it in mermaid using the data below_

# Inputs
 - redcap database tables: redcap_user_information
 - email stream from the sender of automated messages
 - rcc billing database tables: rcc_job_log to get log records about email address rejected by the mail host
 - vivo database tables: WH_UFDIR_PERSON_MDM, WH_GL_ACCT
 
# Outputs
 - redcap database tables: redcap_user_information
 - rcc billing tables: rcc_job_log
