---
title: "create_and_send_new_invoice_line_items"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create_and_send_new_invoice_line_items}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(rcc.billing)
library(DiagrammeR)
```

`create_and_send_new_invoice_line_items` decides what services to bill at the beginning of each month. It reads service request data and redcap project data to create a dataset of records that will be used by the fiscal team to create line items on invoices. The dataset will include records that describe service request work completed in the previous month. The dataset will also have one record for each REDCap project with an anniversary in the previous month and also meets several other criteria.

```{r mermaid-diagram, echo=FALSE, fig.align='center', fig.cap="create_and_send_new_invoice_line_items data processing workflow"}
DiagrammeR::mermaid("
graph TD
    subgraph Service Request Suppliers
        SRS1[REDCap Database]
        SRS2[RCC Billing database]
        SRS3[REDCap Service Request project]
    end
    subgraph Service Request Inputs
        SRI1[redcap_entity_project_ownership]
        SRI2[redcap_projects]
        SRI3[redcap_user_information]
        SRI4[invoice_line_item]
        SRI7[REDCap Service Requests]
    end

    subgraph Project Suppliers
        S1[REDCap Database]
        S2[RCC Billing database]
    end
    subgraph Project Inputs
        I1[redcap_entity_project_ownership]
        I2[redcap_projects]
        I3[redcap_user_information]
        I4[invoice_line_item]
        I5[invoice_line_item_communications]
        I6[service_instance]
    end

    subgraph Process
        P1[Get service request line items]
        P2[Get project line items]
        P3[Combine line items]
        P4[Identify new service instances]
        P5[Combine loggable data]
        P6[Create new invoice_line_item_communications records]
    end
    subgraph Outputs
        O1[invoice_line_item dataset]
        O2[invoice_line_item inserts]
        O3[invoice_line_item_communications inserts]
        O4[service_instance updates]
        O5[Log data]
    end
    subgraph Customers
        C1[RCC Billing System]
        C2[RCC Billing System: rcc_job_log]
        C3[Fiscal team]
    end
    SRS1 --> SRI1
    SRS1 --> SRI2
    SRS1 --> SRI3
    SRS2 --> SRI4
    SRS3 --> SRI7

    S1 --> I1
    S1 --> I2
    S1 --> I3
    S2 --> I4
    S2 --> I5
    S2 --> I6
    
    SRI1 --> P1
    SRI2 --> P1
    SRI3 --> P1
    SRI4 --> P1
    SRI7 --> P1

    I1 --> P2
    I2 --> P2
    I3 --> P2
    I4 --> P2
    
    P1 --> P3
    P2 --> P3
    
    I6 --> P4
    P3 --> P4
    
    I5 --> P6
    P3 --> P6
    
    P3 --> O1
    P3 --> O2
    
    P6 --> O3
    
    P4 --> O4
    
    P3 --> P5
    P4 --> P5
    
    P5 --> O5
    
    O1 --> C3
    O2 --> C1
    O3 --> C1
    O4 --> C1
    O5 --> C2
", width = 850, height= 300)
```
