---
title: "delete_abandoned_projects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{delete_abandoned_projects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(rcc.billing)
library(DiagrammeR)
```

`delete_abandoned_projects` identifies sequestered projects associated with old unpaid invoices and deletes those projects. It also identifies old orphaned projects that have been sequestered for a year and abandons those as well. These deletion operations are equivalent to a normal REDCap deletion event done through the UI. These projects can be undeleted through the UI.

```{r mermaid_diagram, echo=FALSE, fig.align='center'}
DiagrammeR::mermaid("
graph TD
    subgraph Suppliers
        S1[REDCap Database]
        S2[RCC Billing database]
    end
    subgraph Inputs
        I1[redcap_entity_project_ownership]
        I2[redcap_projects]
        I3[invoice_line_item]
        I4[rcc_job_log]
    end
    subgraph Process
        P1[Identify old, unpaid projects]
        P2[Identify old, orphaned projects]
        P3[Identify deleted projects]
        P4[Combine unpaid and<br/>orphaned projects, removing<br/>already-deleted projects]
        P5[Delete the list of projects]
    end
    subgraph Outputs
        O1[deleted_projects dataset]
        O2[redcap_projects updates]
        O3[redcap_log_eventN inserts]
    end
    subgraph Customers
        C1[RCC Billing System: rcc_job_log]
        C2[REDCap Database]
    end
    S1 --> I1
    S1 --> I2
    S2 --> I3
    S2 --> I4
    
    I1 --> P1
    I2 --> P1
    I3 --> P1

    I4 --> P2
    I2 --> P3
    
    P1 --> P4
    P2 --> P4
    P3 --> P4
    P4 --> P5
    P5 --> O1
    P5 --> O2
    P5 --> O3
    O1 --> C1
    O2 --> C2
    O3 --> C2
", width = 700, height= 700)
```
