---
title: "remind_owners_to_review_ownership"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{remind_owners_to_review_ownership}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(rcc.billing)
library(DiagrammeR)
```

`remind_owners_to_review_ownership` sends a quarterly email to every person that owns one or more projects asking them to review the projects they own, reassigning ownership if appropriate, or delete the project if that is appropriate. It provides links to instructions for these changes. The email also explains why this matters--that project owners will be billed annually for each of their projects.

```{r mermaid-diagram, echo=FALSE, fig.align='center', fig.cap="remind_owners_to_review_ownership data processing workflow"}
DiagrammeR::mermaid("
graph TD
    subgraph Suppliers
        S1[REDCap Database]
    end
    subgraph Inputs
        I1[redcap_projects]
        I2[redcap_config]
        I3[redcap_entity_project_ownership]
        I4[redcap_user_information]
    end
    subgraph Process
        P1[Identify a list of non-deleted, non-sequester projects at least 5 months old]
        P2[Group projects by owner]
        P3[Email each person who has projects being sequestered]
    end
    subgraph Outputs
        O1[rcc_job_log]
        O2[Emails about ownership]
    end
    subgraph Customers
        C1[RCC Billing System: rcc_job_log]
        C2[Project owners]
    end
    S1 --> I1
    S1 --> I2
    S1 --> I3
    S1 --> I4

    I1 --> P1
    I3 --> P1

    P1 --> P2
    I4 --> P2
    
    P2 --> P3
    I2 --> P3

    P3 --> O1
    P3 --> O2
    
    O1 --> C1
    O2 --> C2
", width = 700, height= 500)
```
