---
title: "request_correction_of_bad_ownership_data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{request_correction_of_bad_ownership_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(rcc.billing)
library(DiagrammeR)
```

`request_correction_of_bad_ownership_data` sends one of two emails to project designers of every project that has no owner email address. For non-sequestered projects, the email explains projects with no valid email address will be sequestered after 30 days. The email further explains that projects sequestered for one year will be deleted.
The email provides a link to the project ownership page and links to instructions for changing ownership and deleting projects. The script sends the message a few times over the course of a month.

For sequestered projects, the email states the project was sequestered today because it appears to be abandoned. It offers instructions to open a service request if the sequestration is not abandoned and provides the link to submit the service request.

```{r mermaid-diagram, echo=FALSE, fig.align='center', fig.cap="request_correction_of_bad_ownership_data data processing workflow"}
DiagrammeR::mermaid("
graph TD
    subgraph Suppliers
        S1[REDCap Database]
    end
    subgraph Inputs
        I1[redcap_projects]
        I2[redcap_config]
        I3[redcap_entity_project_ownership]
        I4[redcap_user_information]
        I5[redcap_user_rights]
        I6[redcap_user_roles]
    end
    subgraph Process
        P1[Identify projects with unresolvable ownership issues]
        P2[Identify project designers]
        P3[Test sequestration state]
        P4[Email each designer on each project with ownership issues]
        P5[Email each designer about the newly sequestered project]
    end
    subgraph Outputs
        O1[rcc_job_log]
        O2[Emails about ownership issues]
        O3[Emails about today's sequestration]
    end
    subgraph Customers
        C1[RCC Billing System: rcc_job_log]
        C2[Project designers]
    end
    S1 --> I1
    S1 --> I2
    S1 --> I3
    S1 --> I4
    S1 --> I5
    S1 --> I6

    I1 --> P1
    I3 --> P1

    P1 --> P2
    I2 --> P2
    I4 --> P2
    I5 --> P2
    I6 --> P2
    
    P2 --> P3
    P3 --> P4
    P3 --> P5

    P4 --> O1
    P4 --> O2

    P5 --> O1
    P5 --> O3
    
    O1 --> C1
    O2 --> C2
    O3 --> C2
", width = 700, height= 500)
```
