---
title: "update_invoice_line_items_with_invoicing_details"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{update_invoice_line_items_with_invoicing_details}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(rcc.billing)
library(DiagrammeR)
```

`update_invoice_line_items_with_invoicing_details.R` reads downloaded invoicing details provided by the fiscal team and applies them to billing tables and redcap tables. It does these things:

 - Update invoice_line_item records with payment details.
 - Insert invoice_line_item_communication records for each revised invoice_line_item
 - Update service_instance records with new CTSI Study IDs
 - Mirror invoice_line_item data to REDCap table of the same name
 - Update banned_owners table for each PI who left the university

```{r mermaid_diagram, echo=FALSE, fig.width=8, fig.align='center'}
DiagrammeR::mermaid("
graph LR
    subgraph Suppliers
        S1[Fiscal Team]
        S2[REDCap Database]
        S3[REDCap Billing Database]
    end

    subgraph Inputs
        I1[Invoicing Detail File]
        I2[REDCap: invoice_line_item]
        I3[service_instance]
        I4[invoice_line_item]
        I5[invoice_line_item_communications]
        I6[banned_owners]
    end

    subgraph Process
        P2[Align invoice_line_item<br/>with Payment Details]
        P3[New invoice_line_item_communication<br/>Records]
        P4[Identify new CTSI Study<br/>IDs for service_instance]
        P5[Mirror invoice_line_item<br/>to REDCap]
        P6[Identify new<br/>banned_owners]
        P7[Log All Activities]

        P2 --> P3
        P2 --> P4
        P2 --> P5
        P2 --> P6
        P3 --> P7
        P4 --> P7
        P5 --> P7
        P6 --> P7
    end

    subgraph Outputs
        O1[service_instance<br/>updates]
        O2[invoice_line_item<br/>updates]
        O3[invoice_line_item_communications<br/>inserts]
        O4[banned_owners<br/>updates]
        O5[invoice_line_item<br/>updates]
        O6[rcc_job_log]
    end

    subgraph Customers
        C1[REDCap Billing Database]
        C2[REDCap Database]
        C3[RCC Billing System]
    end

    S1 --> I1
    S2 --> I2
    S3 --> I3
    S3 --> I4
    S3 --> I5
    S3 --> I6

    I1 --> P2
    I2 --> P5
    I3 --> P4
    I4 --> P2
    I5 --> P3
    I6 --> P6

    P2 --> O2
    P3 --> O3
    P4 --> O1
    P5 --> O5
    P6 --> O4
    P7 --> O6

    O1 --> C1
    O2 --> C1
    O3 --> C1
    O4 --> C1
    O5 --> C2
    O6 --> C3
", width = 900, height = 500)
```

