---
title: "update_free_support_time_remaining"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{update_free_support_time_remaining}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(rcc.billing)
library(DiagrammeR)
```

`update_free_support_time_remaining` enforces a business rule to provide a free customer service consultation of up to one hour for each REDCap project. It reads the customer service database, groups the data by project ID, and determines which projects have logged hours which should receive hourly rate adjust down to \$0/hour. This marks the records as _pro bono_. Those records will be invoiced like other work but at the \$0/hour rate and the "pro bono" label.

```{r mermaid-diagram, echo=FALSE, fig.align='center', fig.cap="update_free_support_time_remaining data processing workflow"}
DiagrammeR::mermaid("
graph TD
    subgraph Suppliers
        S1[REDCap Service Request project]
        S2[REDCap database]
    end
    subgraph Inputs
        I1[Service Request dataset]
        I2[redcap_projects]
        I3[free_support_time_remaining]
    end

    subgraph Process
        P1[Sum the pro bono time logged for each REDCap Project]
        P2[Add records noting 1 hour of pro bono time for all other projects]
    end
    subgraph Outputs
        O1[free_support_time_remaining updates and inserts]
    end
    subgraph Customers
        C1[REDCap database]
        C2[RCC Billing System: rcc_job_log]
    end
    
    S1 --> I1
    S2 --> I2
    S2 --> I3
    
    I1 --> P1

    P1 --> P2
    I2 --> P2
    
    P2 --> O1
    I3 --> O1
    
    O1 --> C1
    O1 --> C2

", width = 850, height= 500)
```
