---
title: "warn_owners_of_impending_bill"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{warn_owners_of_impending_bill}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(rcc.billing)
library(DiagrammeR)
```

`warn_owners_of_impending_bill` sends an email to project owners who have projects scheduled to be invoiced in the next month. It reminds the owner of the impending bill and the amount of the charge. The email provides guidance about deleting projects and reassigning ownership. It provides links to instructions for each of these tasks. The email describes the [sequestration](novel_concepts.html#sequestration) process that ensues when an invoice is unpaid after 90 days. 

```{r mermaid-diagram, echo=FALSE, fig.align='center', fig.cap="warn_owners_of_impending_bill data processing workflow"}
DiagrammeR::mermaid("
graph TD
    subgraph Suppliers
        S1[REDCap Database]
        S2[RCC Billing Database]
    end
    subgraph Inputs
        I1[redcap_projects]
        I2[redcap_config]
        I3[redcap_entity_project_ownership]
        I4[redcap_user_information]
        I5[redcap_record_counts]
        I6[invoice_line_item]
    end
    subgraph Process
        P1[Identify projects to be billed in the next month]
        P2[Add project owners and details]
        P3[Email each owner about impending billing for their projects]
    end
    subgraph Outputs
        O1[rcc_job_log]
        O2[Emails about impending bills]
    end
    subgraph Customers
        C1[RCC Billing System: rcc_job_log]
        C2[Project owners]
    end
    S1 --> I1
    S1 --> I2
    S1 --> I3
    S1 --> I4
    S1 --> I5
    S2 --> I6

    I1 --> P1
    I3 --> P1
    I6 --> P1

    P1 --> P2
    I2 --> P2
    I4 --> P2
    I5 --> P2
    
    P2 --> P3

    P3 --> O1
    P3 --> O2

    O1 --> C1
    O2 --> C2
", width = 700, height= 500)
```
